### 2020-03-28 paper/cover_letter.tex
Update cover letter

\documentclass{letter}
\usepackage[left=1in,top=0.15in,twoside=false]{geometry}
\usepackage{color,hyperref}
\definecolor{darkblue}{rgb}{0.0,0.0,0.7}
\hypersetup{colorlinks,breaklinks,
            linkcolor=darkblue,urlcolor=darkblue,
            anchorcolor=darkblue,citecolor=darkblue}

% \signature{Agrim Gupta}
\address{From \\
Mohit Madan\\
Department of Electrical  Engineering\\
Indian Institute of Technology Bombay, Mumbai\\
India\\
E-mail: mohitmadan71@gmail.com
\date{\today}
}

\begin{document}
\begin{letter}{To\\
Prof. Wei Zhang\\
Editor-in-Chief\\
IEEE Wireless Communication Letters}
\vspace{0.5in}

\opening{Dear Prof. Wei Zhang,}

Subject: \textbf{Submission of manuscript title ``Scalar Feedback based Joint Time-Frequency Precoder
  interpolation for  MIMO-OFDM Systems'' for potential publication in
  IEEE Wireless Communication Letters (old reference number  WCL2020-0046).}

Thank you for arranging the prompt reviews of our original manuscript
titled ``Scalar Feedback based Joint Time-Frequency Precoder
interpolation for MIMO-OFDM Systems'' (WCL2020-0046) submitted to the
IEEE Wireless Communications Letters, and providing us an opportunity
to resubmit this manuscript with modifications addressing the issues
raised during the reviews. I would like to thank the reviewers for a
careful reading of the manuscript, as well as for the insightful
points that they have raised regarding various technical aspects of
the manuscript. Based on these reviews, we have made extensive
additions and revisions to our submission to address the reviewers
concerns. In particular, we have addressed the key issues pertaining
to the novelty of the method and the use of causal information for
interpolation of the precoders. Please find attached our comments and
actions taken based on the reviews, along with a document indicative
of all the changes made.

The co-authors of this manuscript are
\begin{enumerate}
\item  Agrim Gupta

Department of Electrical  Engineering

Indian Institute of Technology Bombay, Mumbai

India

E-mail: agrim9gupta@gmail.com

\item Prof. Kumar Appaiah

Department of Electrical  Engineering

Indian Institute of Technology Bombay, Mumbai

India

E-mail: akumar@ee.iitb.ac.in

\end{enumerate}

Should you require additional details, please do e-mail me.

Thank you.
\vspace{0.3in}

Sincerely,

Mohit Madan

\end{letter}
\end{document}